Theory about Normal Basis on $F(2^k)$

### 2017-01-24 dissertation/submission/recha.tex
dissertation: after submission_1st revision

\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
\Indm
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
\Indp

  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {BFS Traversal for FSM Reachability}\label{alg:BFS}
\end{algorithm}
\DecMargin{1em}

@@

\begin{Example}

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/MIPS.pdf}
\caption{A 64-bit sequential multiplication hardware design.}
\label{fig:MIPS}}
\end{figure}

Figure \ref{fig:MIPS} depicts a sequential multiplication hardware implementation within a 64-bit 
MIPS. Initially, one multiplicand is preloaded to the lower 64 bits of the product registers. 
Iteratively, the last significant bit (LSB) of current (temporary) product is used as flag to activate the ALU 
to add on the other multiplicand. For each iteration the data in product registers shifts right by 1 bit.
Finally when the most significant bit (MSB) of preloaded multiplicand arrives at the MSB of product registers,
the registers contains the result -- 128 bits product. The behavior can be described by the following algorithm:

@@

\IncMargin{1em}
\begin{algorithm}[H]
\SetAlgoNoLine
\LinesNumbered
\Indm
 \KwIn{Multiplicand $A,B$}
 \KwOut{Product $C$}
\Indp

  Preload $B$ into lower 64-bit of Product Register $P$\;
  \Repeat{64 Repetitions}
  {
  	\If{Last Bit of Product Register $LSB(P)==1$}
  	{
		$P = P_{1/2}+B$\;
	}
	Right shift $P$\;
  }
\Return{$C=P$}
\caption {Sequential multiplication hardware in 64-bit MIPS}\label{alg:MIPS}
\end{algorithm}
\DecMargin{1em}

\vspace{0.2in}

@@

\begin{Example}[Lagrange's interpolation]

\begin{figure}[H]
\centering{
\includegraphics[width=4in]{newfig/Lagrange.pdf}
\caption{Gate-level netlist for Lagrange's interpolation example.}
\label{fig:Lagrange}}
\end{figure}

Assume we are given gate-level netlist shown in Figure \ref{fig:Lagrange}. It can be written as 3 
Boolean equations:
\begin{align*}
z_0 &= a_0 \lor a_1 \lor a_2 \\
z_1 &= a_1 \land \neg a_2 \lor a_0 \land \neg a_1 \land a_2 \\
z_2 &= a_1 \lor \neg a_0 \land a_2
\end{align*}
Define 2 word-level variables $A,Z$ as input and output:
$$A = \{a_2a_1a_0\},~Z = \{z_2z_1z_0\}$$
To convert bit-level to word-level, we need to find mapping $\mathbb{B}^3 \rightarrow \mathbb{B}^3$,
or $\F_{2^3} \rightarrow \F_{2^3}$. The latter one, as mentioned in preliminaries, is a polynomial
function in $\F_{2^3}$.

For each element in $\F_{2^3}$, we write down the truth table as follows:
\begin{table}[H]
\label{tab:truthtable}
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$.}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\end{table}
Now our objective is to abstract a function over finite field $\F_{2^3}$ in word-level variables, {\it i.e.} 
$Z = \Func(A)$. Recall Lagrange's interpolation formula:
\begin{equation}
\label{eqn:Lagrange}
\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
\end{equation}
The geometric meaning of Lagrange's interpolation in real algebra is: given $N$ points with coordinates $(x_i,y_i)$,
they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
written in the form of Equation \ref{eqn:Lagrange}. In this example, although defined in a Galois field instead of 
the real number field, the essential concept of Lagrange's interpolation remains the same. 
We can get 8 points in the affine space:
\begin{align*}
\text{Generic form}&: (a_2\alpha^2+a_1\alpha+a_0,z_2\alpha^2+z_1\alpha+z_0) \gets (A,Z) \\
\text{Point }1&: (0, 0) \gets (000,000) \\
\text{Point }2&: (1,1) \gets (001,001) \\
\text{Point }3&:  (\alpha,\alpha^2 + \alpha + 1) \gets (010,111)\\
\text{Point }4&:  (\alpha + 1,\alpha^2 + \alpha + 1)\gets (011,111) \\
\text{Point }5&:  (\alpha^2,\alpha^2+1) \gets (100,101)\\
\text{Point }6&:  (\alpha^2 + 1,\alpha+1)\gets (101,011)\\
\text{Point }7&:  (\alpha^2 + \alpha,\alpha^2 + 1)\gets (110,101)\\
\text{Point }8&:  (\alpha^2 + \alpha + 1,\alpha^2 + 1)\gets (111,101)
\end{align*}
Substitute 8 $(x_i,y_i)$ pairs in Equation \ref{eqn:Lagrange} with these 8 points in $\F_{2^3}$.
The result is a polynomial function with degree no greater than 7:
\begin{align*}
Z =& \Func(A) \\
=& (\alpha^2+\alpha+1) A^7 + (\alpha^2+1) A^6 + \alpha A^5 + (\alpha+1) A^4 \\
&+ (\alpha^2+\alpha+1)A^3+ (\alpha^2+1)A
\end{align*}

@@

\subsection{FSM Model for Sequential Circuits}
A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 
circuits. If a FSM's primary outputs depend on primary inputs and present state inputs, it is named as a \textit{Mealy machine};
the formal definition is as follows:
\begin{Definition}
A Mealy machine is an $n$-tuple $\mathcal M = (\Sigma,O,S,S^0,\Delta,\Lambda)$ where
\begin{itemize}
\item $\Sigma$ is the input label, $O$ is the output label;
\item $S$ is the set of states, $S^0\subseteq S$ is the set of initial states;
\item $\Delta:\ S\times\Sigma\to S$ is the next state transition function;
\item $\Lambda:\ S\times\Sigma\to O$ is the output function.
\end{itemize}
\end{Definition}
The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
its primary outputs only depend on the present states, {\it i.e.} the output function is defined as
$$\Lambda:\ S \to O$$
Typical sequential circuits can be depicted as Figure \ref{fig:seqmodel}(a). Primary inputs
$x_1,\dots,x_m \in \Sigma$, and primary outputs $z_1,\dots,z_n\in O$. Signals $s_1,\dots,s_k$ 
are present state (PS) variables, $t_1,\dots,t_k$ are next state (NS) variables.
We can define 2 $k$-bit words denoting the PS/NS variables as there are $k$ flip-flops
in the datapath: $S = (s_1,\dots,s_k), ~T=(t_1,\dots,t_k)$. Transition function
at bit level are defined as $\Delta_i$: 
$$t_i = \Delta_i(s_1,\dots,s_k,x_1,\dots,x_m)$$
\begin{figure}[H]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=5.5in]{newfig/seqmodel.pdf}
% \vspace{-0.2in}
\caption{FSM models of sequential circuits.}
%\end{minipage}
\label{fig:seqmodel}}
\end{figure}
In some cases, arithmetic computations are implemented as Moore machines where input operands
are loaded into register files $R$ and the FSM is executed for $k$ clock cycles.
We can simplify them to the model in Figure \ref{fig:seqmodel}(b).

@@

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/new_stg.eps}
\caption{The example FSM and the gate-level implementation. }
\label{fig:fsm}}
\end{figure}

@@

\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
\Indm
 \KwIn{The circuit's characteristic polynomial ideal $J_{ckt}$,
   initial state polynomial $\F(S)$, and LEX term order: bit-level
   variables $x, s, t$ $>$ PS word $S$ $>$ NS word $T$}
\Indp

  $from^0 = reached = \Func(S)$\;
  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
  {
  	$i \gets i + 1$\;
  	$G \gets$GB($\langle J_{ckt},J_v, from^{i-1}\rangle$)\;
  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest}
	$to^i \gets G\cap \Fkk[T]$\;
	\tcc{There will be a univariate polynomial in $G$ denoting the
          set of next states in word-level variable $T$}
	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
  	\tcc{Use product of ideals to attain a union of newly reached states and formerly reached states}
	$from^i \gets new^i(S\setminus T)$\;
	\tcc{Start a new iteration by replacing variable $T$ in newly reached states with current state variable $S$}
  }
  \tcc{Loop until a fixpoint reached: newly reached state is empty}
\Return{$\langle reached\rangle$}
\caption {Algebraic Geometry based FSM Traversal}\label{alg:univa}
\end{algorithm}
\DecMargin{1em}

@@

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/proj_reacha.pdf}
\caption{Projection of the variety of circuit description ideal.}
\label{fig:proj_reacha}}
\end{figure}

@@

\IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
\Indm
 \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\Func(S)$}
 \KwOut{Final reachable states represented by polynomial $\mathcal G(T)$}
\Indp

  $from^0 = reached = \Func(S)$\;
  $f_T = $Reduce($Spoly(f_w,f_g), J_{ckt})$\;
  \tcc{Compute $Spoly$ for the critical pair, then reduce it with circuit ideal under RATO}
	Eliminate bit-level variables in $f_T$\;
  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
  {
  	$i \gets i + 1$\;
  	$G \gets$GB($\langle f_T , from^{i-1}\rangle+J_0^{PI}$)\;
  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest; $J_0^{PI}$ covers all possible inputs from PIs}
	$to^i \gets G\cup \Fkk[T]$\;
	\tcc{There will be a univariate polynomial in $G$ denoting next state in word-level variable $T$}
	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
  	\tcc{Use product of ideals to attain a union of new reached states and formerly reached states}
	$from^i \gets new^i(S\setminus T)$\;
	\tcc{Start a new iteration by replacing variable $T$ in new reached states with current state variable $S$}
  }
  \tcc{Loop until a fix-point reached: newly reached state is empty}
\Return{$\langle reached\rangle$}
\caption {Refined Algebraic Geometry based FSM Traversal}\label{alg:refined}
\end{algorithm}
\DecMargin{1em}

@@

@@

Assume we have $k$-bit datapath and $n$-bit PIs. In finite field, we need to carefully partition $n$ PIs
such that states of each partition can be covered by a univariate polynomial respectively.

@@

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/flowchart.pdf}
\caption{Execution process of word-level FSM traversal tool.}
\label{fig:flowchart}}
\end{figure}

@@

After the script is executed, the traversal finishes after 4 
transition iterations, which denotes BFS depth equals to 4. The final reachable states 
is a degree-9 polynomial in $T$, indicating final reachable states set contains 9 states.
And state encodings can be obtained by solving this polynomial equation.

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/SING.pdf}
\caption{Singular script for executing bit-to-word substitution and traversal loop.}
\label{fig:SING}}
\end{figure}

\begin{figure}[H]
\centering{
\includegraphics[width=\textwidth]{newfig/Recha_result.pdf}
\caption{The output given by our traversal tool.}
\label{fig:Recha_result}}
\end{figure}

@@

\begin{table}[H]
\centering
\caption{Results of running benchmarks using our tool. 
\small{Parts I to III denote the time taken by polynomial divisions,
  bit-level to word-level abstraction and iterative reachability
  convergence checking part of our approach, respectively.}}
{\small 
\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|}
\hline
\multirow{3}{*}{\centering Benchmark} 
& \multirow{3}{0.9cm}{\centering \# Gates} 
& \multirow{3}{1.1cm}{\centering \# Latches} 
& \multirow{3}{*}{\centering \# PIs}
 & \multirow{3}{0.9cm}{\centering \# States}
 & \multirow{3}{1.2cm}{\centering \# iterations}
 & \multicolumn{3}{c|}{\multirow{2}{2.0cm}{\centering Runtime (sec)}}
 & \multirow{3}{1.3cm}{\centering Runtime of VIS (sec)} \\
  & & & & & &\multicolumn{3}{c|}{}& \\
  \cline{7-9}
    & & & & & & I & II & III & \\
\hline
\hline
b01 & 39  & 5  & 2 & 18  & 5  & $<0.01$ & 0.01 & 0.02 & $<0.01$\\
b02 & 24  & 4  & 1 & 8 & 5 & $<0.01$  & 0.01 & $<0.01$ & $<0.01$ \\
b06 & 49  & 9  & 2 & 13 & 4 & $<0.01$ & 0.07 & 5.0 & $<0.01$ \\
s27 & 10 & 3 & 4 & 6 & 2 & $<0.01$ & 0.01 & 0.02  & $<0.01$  \\
s208 & 61 & 8 & 11 & 16 & 16 & $<0.01$ & 0.32 & 2.4 & $<0.01$ \\
s386 & 118 & 6 & 13 & 13  & 3 & 1.0 & 7.6 & 8.2  & $<0.01$ \\
bbara & 82 & 4 & 4 & 10 & 6 & 0.04 & 0.01 & 0.04  & $<0.01$ \\
beecount & 48  & 3  & 3 & 7  & 3  &$<0.01$ & 0.01 & 0.01 & $<0.01$ \\
dk14 & 120  & 3  & 3 & 7  & 2  & 45 & $<0.01$ & 0.08 & $<0.01$\\
donfile & 205  & 5  & 2 & 24 & 3  & 12316 & 0.02 & 1.7  & $<0.01$\\
\hline
\end{tabular}
}
\label{tab:recha_result}  
\end{table}

@@

%\vspace{-0.1in}

>>>>

% \IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
% \Indm
 \KwIn{Transition functions $\Delta$, initial state $S^0$}
% \Indp

  $from^0 = reached = S^0$\;
  \Repeat{$new^i == 0$}
  {
  	$i \gets i + 1$\;
	$to^i \gets$Img$(\Delta, from^{i-1})$\;
	$new^i \gets to^i \cap \overline{reached}$\;
  	$reached \gets reached \cup new^i$\;
	$from^i \gets new^i$\;
  }
\Return{$reached$}
\caption {BFS Traversal for FSM Reachability}\label{alg:BFS}
\end{algorithm}
% \DecMargin{1em}

@@

\begin{Example}
Figure \ref{fig:MIPS} depicts a sequential multiplication hardware implementation within a 64-bit 
MIPS. Initially, one multiplicand is preloaded to the lower 64 bits of the product registers. 
Iteratively, the last significant bit (LSB) of current (temporary) product is used as flag to activate the ALU 
to add on the other multiplicand. For each iteration the data in product registers shifts right by 1 bit.
Finally when the most significant bit (MSB) of preloaded multiplicand arrives at the MSB of product registers,
the registers contains the result -- 128 bits product. The behavior can be described by the following algorithm:

@@

% \IncMargin{1em}
\begin{algorithm}[H]
\SetAlgoNoLine
\LinesNumbered
% \Indm
 \KwIn{Multiplicand $A,B$}
 \KwOut{Product $C$}
% \Indp

  Preload $B$ into lower 64-bit of Product Register $P$\;
  \Repeat{64 Repetitions}
  {
  	\If{Last Bit of Product Register $LSB(P)==1$}
  	{
		$P = P_{1/2}+B$\;
	}
	Right shift $P$\;
  }
\Return{$C=P$}
\caption {Sequential multiplication hardware in 64-bit MIPS}\label{alg:MIPS}
\end{algorithm}
% \DecMargin{1em}

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/MIPS.pdf}
\caption{A 64-bit sequential multiplication hardware design.}
\label{fig:MIPS}}
\end{figure}

@@

\begin{Example}[Lagrange's interpolation]
Assume we are given gate-level netlist shown in Figure \ref{fig:Lagrange}. It can be written as 3 
Boolean equations:
\begin{align*}
z_0 &= a_0 \lor a_1 \lor a_2 \\
z_1 &= a_1 \land \neg a_2 \lor a_0 \land \neg a_1 \land a_2 \\
z_2 &= a_1 \lor \neg a_0 \land a_2
\end{align*}
Define 2 word-level variables $A,Z$ as input and output:
$$A = \{a_2a_1a_0\},~Z = \{z_2z_1z_0\}$$
To convert bit-level to word-level, we need to find mapping $\mathbb{B}^3 \rightarrow \mathbb{B}^3$,
or $\F_{2^3} \rightarrow \F_{2^3}$. The latter one, as mentioned in preliminaries, is a polynomial
function in $\F_{2^3}$.

\begin{figure}[bp]
\centering{
\includegraphics[width=4in]{newfig/Lagrange.pdf}
\caption{Gate-level netlist for Lagrange's interpolation example.}
\label{fig:Lagrange}}
\end{figure}

For each element in $\F_{2^3}$, we write down the truth table as Table \ref{tab:truthtable}.
Now our objective is to abstract a function over finite field $\F_{2^3}$ in word-level variables, {\it i.e.} 
$Z = \Func(A)$. Recall Lagrange's interpolation formula:
\begin{equation}
\label{eqn:Lagrange}
\Func(x) =  \sum_{k=1}^N \left[ \prod_{(0\leq j \leq k-1),(j\neq i)}\frac{x-x_j}{x_i-x_j} \cdot y_k \right]
\end{equation}

\begin{table}[bp]
\caption{Truth table for mappings in $\mathbb{B}^3$ and $\F_{2^3}$.}
\label{tab:truthtable}
\centering{\small
\begin{tabular}{|c|ccc|c|}
\hline
$\{a_2a_1a_0\}\in\mathbb{B}^3$   & $A\in \F_{2^3}$ &$\rightarrow$& $\{z_2z_1z_0\}\in\mathbb{B}^3$  & $Z\in \F_{2^3}$ \\
\hline
000  &0 &$\rightarrow$&000 & 0 \\
001  &1 &$\rightarrow$&001 & 1 \\
010  &$\alpha$ & $\rightarrow$ & 111& $\alpha^2 + \alpha + 1$ \\
011  &$\alpha + 1$ &$\rightarrow$& 111 & $\alpha^2 + \alpha + 1$ \\
100  &$\alpha^2$ &$\rightarrow$& 101 &  $\alpha^2+1$ \\
101  &$\alpha^2 + 1$ &$\rightarrow$&011 & $\alpha+1$ \\
110  &$\alpha^2 + \alpha$&$\rightarrow$& 101 &$\alpha^2 + 1$ \\
111  &$\alpha^2 + \alpha + 1$ &$\rightarrow$& 101 &$\alpha^2 + 1$\\
\hline
\end {tabular}
}
\end{table}

The geometric meaning of Lagrange's interpolation in real algebra is: given $N$ points with coordinates $(x_i,y_i)$,
they can always be fitted into a polynomial function with at most $N-1$ degree, and that function can be 
written in the form of Equation \ref{eqn:Lagrange}. In this example, although defined in a Galois field instead of 
the real number field, the essential concept of Lagrange's interpolation remains the same. 
We can get 8 points in the affine space:
\begin{align*}
\text{Generic form}&: (a_2\alpha^2+a_1\alpha+a_0,z_2\alpha^2+z_1\alpha+z_0) \gets (A,Z) \\
\text{Point }1&: (0, 0) \gets (000,000) \\
\text{Point }2&: (1,1) \gets (001,001) \\
\text{Point }3&:  (\alpha,\alpha^2 + \alpha + 1) \gets (010,111)\\
\text{Point }4&:  (\alpha + 1,\alpha^2 + \alpha + 1)\gets (011,111) \\
\text{Point }5&:  (\alpha^2,\alpha^2+1) \gets (100,101)\\
\text{Point }6&:  (\alpha^2 + 1,\alpha+1)\gets (101,011)\\
\text{Point }7&:  (\alpha^2 + \alpha,\alpha^2 + 1)\gets (110,101)\\
\text{Point }8&:  (\alpha^2 + \alpha + 1,\alpha^2 + 1)\gets (111,101)
\end{align*}
Substitute 8 $(x_i,y_i)$ pairs in Equation \ref{eqn:Lagrange} with these 8 points in $\F_{2^3}$.
The result is a polynomial function with degree no greater than 7:
\begin{align*}
Z =& \Func(A) \\
=& (\alpha^2+\alpha+1) A^7 + (\alpha^2+1) A^6 + \alpha A^5 + (\alpha+1) A^4 \\
&+ (\alpha^2+\alpha+1)A^3+ (\alpha^2+1)A
\end{align*}

@@

\subsection{FSM Model for Sequential Circuits}
A finite state machine (FSM) is a mathematical model of computation for designing and analyzing sequential logic 
circuits. If a FSM's primary outputs depend on primary inputs and present state inputs, it is named as a \textit{Mealy machine};
the formal definition is as follows:
\begin{Definition}
A Mealy machine is an $n$-tuple $\mathcal M = (\Sigma,O,S,S^0,\Delta,\Lambda)$ where
\begin{itemize}
\item $\Sigma$ is the input label, $O$ is the output label;
\item $S$ is the set of states, $S^0\subseteq S$ is the set of initial states;
\item $\Delta:\ S\times\Sigma\to S$ is the next state transition function;
\item $\Lambda:\ S\times\Sigma\to O$ is the output function.
\end{itemize}
\end{Definition}
The other kind of FSM is \textit{Moore machine}, its difference from Mealy machine is that
its primary outputs only depend on the present states, {\it i.e.} the output function is defined as
$$\Lambda:\ S \to O$$
Typical sequential circuits can be depicted as Figure \ref{fig:seqmodel}(a). Primary inputs
$x_1,\dots,x_m \in \Sigma$, and primary outputs $z_1,\dots,z_n\in O$. Signals $s_1,\dots,s_k$ 
are present state (PS) variables, $t_1,\dots,t_k$ are next state (NS) variables.
We can define 2 $k$-bit words denoting the PS/NS variables as there are $k$ flip-flops
in the datapath: $S = (s_1,\dots,s_k), ~T=(t_1,\dots,t_k)$. Transition function
at bit level are defined as $\Delta_i$: 
$$t_i = \Delta_i(s_1,\dots,s_k,x_1,\dots,x_m)$$

In some cases, arithmetic computations are implemented as Moore machines where input operands
are loaded into register files $R$ and the FSM is executed for $k$ clock cycles.
We can simplify them to the model in Figure \ref{fig:seqmodel}(b).

\begin{figure}[bp]
\centering{
%\begin{minipage}{12cm}
\includegraphics[width=5.5in]{newfig/seqmodel.pdf}
% \vspace{-0.2in}
\caption{FSM models of sequential circuits.}
%\end{minipage}
\label{fig:seqmodel}}
\end{figure}

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/new_stg.eps}
\caption{The example FSM and the gate-level implementation. }
\label{fig:fsm}}
\end{figure}

@@

% \IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
% \Indm
 \KwIn{The circuit's characteristic polynomial ideal $J_{ckt}$,
   initial state polynomial $\F(S)$, and LEX term order: bit-level
   variables $x, s, t$ $>$ PS word $S$ $>$ NS word $T$}
% \Indp

  $from^0 = reached = \Func(S)$\;
  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
  {
  	$i \gets i + 1$\;
  	$G \gets$GB($\langle J_{ckt},J_v, from^{i-1}\rangle$)\;
  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest}
	$to^i \gets G\cap \Fkk[T]$\;
	\tcc{There will be a univariate polynomial in $G$ denoting the
          set of next states in word-level variable $T$}
	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
  	\tcc{Use product of ideals to attain a union of newly reached states and formerly reached states}
	$from^i \gets new^i(S\setminus T)$\;
	\tcc{Start a new iteration by replacing variable $T$ in newly reached states with current state variable $S$}
  }
  \tcc{Loop until a fixpoint reached: newly reached state is empty}
\Return{$\langle reached\rangle$}
\caption {Algebraic Geometry based FSM Traversal}\label{alg:univa}
\end{algorithm}
% \DecMargin{1em}

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/proj_reacha.pdf}
\caption{Projection of the variety of circuit description ideal.}
\label{fig:proj_reacha}}
\end{figure}

@@

% \IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
% \Indm
 \KwIn{Input-output circuit characteristic polynomial ideal $J_{ckt}$, initial state polynomial $\Func(S)$}
 \KwOut{Final reachable states represented by polynomial $\mathcal G(T)$}
% \Indp

  $from^0 = reached = \Func(S)$\;
  $f_T = $Reduce($Spoly(f_w,f_g), J_{ckt})$\;
  \tcc{Compute $Spoly$ for the critical pair, then reduce it with circuit ideal under RATO}
	Eliminate bit-level variables in $f_T$\;
  \Repeat{$\langle new^i\rangle == \langle 1\rangle$}
  {
  	$i \gets i + 1$\;
  	$G \gets$GB($\langle f_T , from^{i-1}\rangle+J_0^{PI}$)\;
  	\tcc{Compute Gr\"obner basis with elimination term order: $T$ smallest; $J_0^{PI}$ covers all possible inputs from PIs}
	$to^i \gets G\cup \Fkk[T]$\;
	\tcc{There will be a univariate polynomial in $G$ denoting next state in word-level variable $T$}
	$\langle new^i\rangle \gets \langle to^i\rangle + (\langle T^{2^k}-T\rangle:\langle reached\rangle)$\;
	\tcc{Use quotient of ideals to attain complement of reached states, then use sum of ideals to attain an intersection with next state}
  	$\langle reached\rangle \gets \langle reached\rangle \cdot \langle new^i\rangle$\;
  	\tcc{Use product of ideals to attain a union of new reached states and formerly reached states}
	$from^i \gets new^i(S\setminus T)$\;
	\tcc{Start a new iteration by replacing variable $T$ in new reached states with current state variable $S$}
  }
  \tcc{Loop until a fix-point reached: newly reached state is empty}
\Return{$\langle reached\rangle$}
\caption {Refined Algebraic Geometry based FSM Traversal}\label{alg:refined}
\end{algorithm}
% \DecMargin{1em}

@@

As in Figure \ref{fig:PI}, assume there exist $k$-bit datapath and $n$-bit PIs. In finite field, we need to carefully partition $n$ PIs
such that states of each partition can be covered by a univariate polynomial respectively.

@@

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/flowchart.pdf}
\caption{Execution process of word-level FSM traversal tool.}
\label{fig:flowchart}}
\end{figure}

@@

After the script is executed, the traversal finishes after 4 
transition iterations, which denotes BFS depth equals to 4. According to the output in
Figure \ref{fig:Recha_result}, the final reachable states 
is a degree-9 polynomial in $T$, indicating final reachable states set contains 9 states.
And state encodings can be obtained by solving this polynomial equation.

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/SING.pdf}
\caption{Singular script for executing bit-to-word substitution and traversal loop.}
\label{fig:SING}}
\end{figure}

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/Recha_result.pdf}
\caption{The output given by our traversal tool.}
\label{fig:Recha_result}}
\end{figure}

@@

@@

\begin{table}[tbp]
\centering
\caption{Results of running benchmarks using our tool. 
\small{Parts I to III denote the time taken by polynomial divisions,
  bit-level to word-level abstraction and iterative reachability
  convergence checking part of our approach, respectively.}}
{\small 
\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|}
\hline
\multirow{3}{*}{\centering Benchmark} 
& \multirow{3}{0.9cm}{\centering \# Gates} 
& \multirow{3}{1.1cm}{\centering \# Latches} 
& \multirow{3}{*}{\centering \# PIs}
 & \multirow{3}{0.9cm}{\centering \# States}
 & \multirow{3}{1.2cm}{\centering \# iterations}
 & \multicolumn{3}{c|}{\multirow{2}{2.0cm}{\centering Runtime (sec)}}
 & \multirow{3}{1.3cm}{\centering Runtime of VIS (sec)} \\
  & & & & & &\multicolumn{3}{c|}{}& \\
  \cline{7-9}
    & & & & & & I & II & III & \\
\hline
\hline
b01 & 39  & 5  & 2 & 18  & 5  & $<0.01$ & 0.01 & 0.02 & $<0.01$\\
b02 & 24  & 4  & 1 & 8 & 5 & $<0.01$  & 0.01 & $<0.01$ & $<0.01$ \\
b06 & 49  & 9  & 2 & 13 & 4 & $<0.01$ & 0.07 & 5.0 & $<0.01$ \\
s27 & 10 & 3 & 4 & 6 & 2 & $<0.01$ & 0.01 & 0.02  & $<0.01$  \\
s208 & 61 & 8 & 11 & 16 & 16 & $<0.01$ & 0.32 & 2.4 & $<0.01$ \\
s386 & 118 & 6 & 13 & 13  & 3 & 1.0 & 7.6 & 8.2  & $<0.01$ \\
bbara & 82 & 4 & 4 & 10 & 6 & 0.04 & 0.01 & 0.04  & $<0.01$ \\
beecount & 48  & 3  & 3 & 7  & 3  &$<0.01$ & 0.01 & 0.01 & $<0.01$ \\
dk14 & 120  & 3  & 3 & 7  & 2  & 45 & $<0.01$ & 0.08 & $<0.01$\\
donfile & 205  & 5  & 2 & 24 & 3  & 12316 & 0.02 & 1.7  & $<0.01$\\
\hline
\end{tabular}
}
\label{tab:recha_result}  
\end{table}
